Structured Genealogy Processes

### 2022-02-01 ms.Rnw
DSABNS, Bilbao, 2022-02-11

\begin{figure}
  \resizebox{0.6\linewidth}{!}{
    \begin{tikzpicture}[scale=1.8]
      \usetikzlibrary{arrows}
      \tikzstyle{coordinate}=[inner sep=0pt,outer sep=0pt]
      \tikzstyle{det}=[color=black, thick, >=stealth]
      \tikzstyle{stoch}=[color=black, dashed, thick, >=stealth]
      \node (A) at (0,0) {$\Master_t$};
      \node (H) at (30:1) {$\Hist_t$};
      \node (X) at (30:2) {$\X_t$};
      \node (G) at (-30:1) {$\Gen_t$};
      \node (I) at (-30:2) {$\Inv_t$};
      \node (P) at (0:1.732) {$\Prune_t$};
      \node (O) at (1.732,.5) {$\Och_i$};
      \node (V) at (0:2.732) {$\Vis_t$};
      \node (W) at (2.732,.5) {$\Wch_i$};
      \draw [det,->] (A) -- (H);
      \draw [det,->] (H) -- (X);
      \draw [det,->] (A) -- (G);
      \draw [det,->] (G) -- (I) node[midway,below,sloped] {$\mathrm{ext}$};
      \draw [det,->] (G) -- (P) node[midway,below,sloped] {$\mathrm{prune}$};
      \draw [det,->] (P) -- (V) node[midway,below,sloped] {$\mathrm{obs}$};
      \draw [det,->] (P) -- (O);
      \draw [det,->] (V) -- (W);
      \draw [stoch,->] (H) -- (G);
    \end{tikzpicture}
  }
  \caption{
    \label{fig:constellation}
    Relations among the various Markov processes discussed in the paper.
    Deterministic maps are indicated with solid arrows;
    random maps are shown as dashed arrows.
    All the maps shown commute.
    $\X_t$ is the \emph{population process}, a model of the dynamics of some system, which we take as a starting point.
    $\Hist_t$ is the \emph{history process}, which records the full history of $\X_t$.
    $\Inv_t$ is the \emph{inventory process}: at each time $t$, $\Inv_t$ is an inventory of all extant individuals in the population, each of which has a globally unique name.
    $\Gen_t$ is the \emph{genealogy process}, which captures the precise genealogical relationships among all individuals in $\Inv_t$, as well as among any samples that have been taken from the population.
    $\Vis_t$ is the \emph{visible genealogy process}, which is $\Gen_t$ pruned so that only relationships among samples remain.
    Finally $\Wch_i$ is the \emph{embedded chain of the visible genealogy process}, which is $\Vis_{s_i}$, $s_i$ being the time of the $i$-th sample.
    All of these processes can be obtained via deterministic procedures applied to the \emph{master process} $\Master_t$, as described in the text.
  }
\end{figure}

@@

\section{Likelihood, filtering equation, and algorithms}

\section{Results}
\label{sec:results}
The derivation of the probability density of $\Wch_i$ given $\Hist_{S_i}$ can be divided into two steps:
\begin{align}
  P_{\Wch_i|\Hist}&=\sum_{\Och_i}\,P_{\Wch_i|\Och_i}P_{\Och_i|\Hist} \\
                  &=\sum_{\Och_i}\,P_{\Wch_i|\Och_i}\left(P_{\Och_1|\Hist}\prod_{q=2}^{i}P_{\Och_q|\Och_{q-1}}\right) 
\end{align}
% Trivially, $P_{\Och_1|\Hist}=1$ and $P_{\Wch_1|\Och_1}=1$ such that $P_{\Wch_1|\Hist}=1$.

@@

We define $\phi_{qk}\coloneqq\sum_{\omega\in\Omega_q}\auxmeas_{u_k,x_k-u_k}(N_k(\omega)) = \prod_{k=1}^K\sum_{\omega\in\Omega_q}\left(\sum_{j=1}^J\dfrac{\indicator{\Inv_{t_{k-1}}^j(\omega)}(N_k(\omega))}{I^j(x_k-u_k)}\right)$.

@@

All cases can be categorized into four groups based on the occurrence of population event within time $\halfopen{a_q,s_q}$: no population event or simultaneous events, a birth from Deme $j$ to $j'$, a sample of Deme $j'$, and a migration from Deme $j$ to $j'$ where $j\neq j'$. Within each group, we investigate $\phi_{qk}^{j''}$ by considering whether the corresponding population event was presented in the pruned genealogy and involving the $q$-th lineage sample, which was in Deme $j''$. For $q>1$, we then have:

\begin{enumerate}[(a)]
  \item If $t_k\notin\halfopen{a_q,s_q}$, or if $u_k\notin\Bset\cup\Sset\cup\Mset$, so we have $\phi_{qk}^{j''}=1$.
  \item If $t_k\in\Cset(\Och_{q-1})\cup\Qset(\Och_{q-1})$, then again, $\phi_{qk}^{j''}=1$.
  \item If $j''\notin\{j,j'\}$, then the new lineage sample was not involved in any population events involving Demes $j$ and $j'$ at time $t_k$ and $\phi_{qk}^{j''}=1$.
  \item \label{item:M2Z} If $t_k\in(a_q,s_q)\setminus\Masterset(\Och_{q-1})$, $t_k\in\Zset^{j,j'}(\Och_q)$, and $u_k\in\Mset^{j,j'}$, then there was a migration event at time $t_k$ and the $j$-th sample lineage was involved. 
  Whether this migration event was presented as a deme-transition event in $\Och_q$ depends on whether $j''=j'$.
  The population size of Deme $j$ right before time $t_k$ was $I_{k-1}^j$, while that of Deme $j'$ right after $t_k$ was $I_k^{j'}$.
  Lineages of $\Och_{q-1}$ were not involved in this migration event, so they were excluded.
  Thus, at time $t_k$, one of the $I_{k-1}^{j}-\ell_{q,k-1}^{j}$ individuals was chosen to migrate and right after, among $I_{k}^{j'}-\ell_{qk}^{j'}$ unobserved individuals, the $j$-th sample was the exact one just migrated.
  Therefore, $\phi_{qk}^{j''}=\delta_{j',j''}/\left((I_{k-1}^{j}-\ell_{q,k-1}^{j})(I_{k}^{j'}-\ell_{qk}^{j'})\right)$.
  \item If $t_k\in(a_q,s_q),\ t_k\notin\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Mset^{j,j'}$, the logic is the same as that in case \ref{item:M2Z}, but the migration event was not presented in $\Och_q$. One can avoid the exact genealogical deme-transition induced by a migration at time $t_k$, with $\phi_{qk}^{j''}=1-\delta_{j',j''}/\left((I_{k-1}^{j}-\ell_{q,k-1}^{j})(I_{k}^{j'}-\ell_{qk}^{j'})\right)$.
  \item \label{item:G2D} If $t_k=a_q\in\Lset^{j}(\Och_{q-1})$, and obviously, $u_k\in\Sset_j$.
    A individual in Deme $j$ was sampled at $t_k$ and the $q$-th lineage sample was exactly the direct-descend of it, \ie a direct-descend event caused by the $q$-th lineage sample was presented in $\Och_q$ at time $t_k$.

  \item
  \item If $t_k\in(a_q,s_q)\setminus\Masterset(\Och_{q-1})$, $t_k\in\Zset^{j,j'}(\Och_q)$, and $u_k\in\Bset^{j,j'}$, then there was a birth event at time $t_k$ and the $j$-th sample lineage was involved. 
    However, this birth event was observed as a deme-transition event (\ie $j\neq j'$) in $\Och_q$, instead of a coalescent event. 
    This birth event resulted in a newly added node holding two black balls, which were separately from Deme $j$ and Deme $j'$.
    Thus, the population size of Deme $j$ and $j'$, right after $t_k$, were $I_k^j$ and $I_k^{j'}$, respectively. 
    There were in total $(I_k^{j}-\ell_{qk}^{j})(I_k^{j'}-\ell_{qk}^{j'})$ combinations of the black balls, because none of the lineages in $\Och_{q-1}$ was involved. 
    Among all these combinations, the $j$-th sample lineage was exactly the black ball newly added to Deme $j'$ after the birth event.
    Therefore, $\phi_{qk}^{j''}=\delta_{j',j''}/\left((I_k^{j}-\ell_{qk}^{j})(I_k^{j'}-\ell_{qk}^{j'})\right)$.
  \item
  \item
\end{enumerate}

@@

    % deme-transition time can evolve to coalescent time, when the new sample is the parent of a "cross-deme" birth
    1, &\text{if}\ t_k\in\Cset(\Och_{q-1})\cup\Qset(\Och_{q-1}),\\[12pt]
    1, &\text{if}\ j''\notin\{j,j'\}\\[12pt]
    \frac{\delta_{j',j''}}{(I_{k-1}^j-\ell_{q,k-1}^j)(I_k^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\in\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Mset^{j,j'},\\[12pt]
    1-\frac{\delta_{j',j''}}{(I_{k-1}^j-\ell_{q,k-1}^{j})(I_k^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\notin\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Mset^{j,j'},\\[12pt]
    \frac{\delta_{j',j''}}{I_k^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k=a_q\in\Lset^{j'}(\Och_{q-1}),\\[12pt]
    1-\frac{\delta_{j',j''}}{I_k^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k\in(a_q,s_q)\setminus\Qset(\Och_{q-1})\ \text{and}\ u_k\in\Sset^{j'},\\[12pt]
    \frac{\delta_{j',j''}}{(I_k^j-\ell_{qk}^j)(I_k^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\in\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Bset^{j,j'},\\[12pt]

    % B2C: either j''=j' or j''=j, the numerator = 1 when j'!=j and =2 when j'=j.
    \frac{1+\delta_{j,j'}}{I_k^{j}(I_k^{j'}-\delta_{j,j'})-\ell_{qk}^j(\ell_{qk}^{j'}-\delta_{j,j'})}, &\text{if}\ t_k=a_q\in\Cset^{j,j'}(\Och_q),\\[12pt]
    1-\frac{(1+\delta_{j,j'})\ell_{qk}^{j}}{I_k^{j}(I_k^{j'}-\delta_{j,j'})-\ell_{qk}^j(\ell_{qk}^{j'}-\delta_{j,j'})}-\frac{1}{I_k^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\notin\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Bset^{j,j'},\\[12pt]
    0, &\text{otherwise}.
  \end{cases}
\end{equation}

@@

\begin{thm}\label{thm:prunehthm1}
  With the definitions as above, $P_{\Och_i|\Hist}(w|h) = \prod_{q=1}^{i}\prod_{k=1}^{K}\phi_{qk}$.
\end{thm}

>>>>

\begin{figure}
  \resizebox{0.6\linewidth}{!}{
    \begin{tikzpicture}[scale=2.5]
      \usetikzlibrary{arrows}
      \tikzstyle{coordinate}=[inner sep=0pt,outer sep=0pt]
      \tikzstyle{det}=[color=black, thick, >=stealth]
      \tikzstyle{stoch}=[color=black, dashed, thick, >=stealth]
      \node (A) at (0,0) {$\Master_t$};
      \node (H) at (30:1) {$\Hist_t$};
      \node (X) at (30:2) {$\X_t$};
      \node (G) at (-30:1) {$\Gen_t$};
      \node (I) at (-30:2) {$\Inv_t$};
      \node (P) at (0:1.732) {$\Prune_t$};
      \node (O) at (1.732,.5) {$\Och_i$};
      \node (V) at (0:2.732) {$\Vis_t$};
      \node (W) at (2.732,.5) {$\Wch_i$};
      \draw [det,->] (A) -- (H);
      \draw [det,->] (H) -- (X);
      \draw [det,->] (A) -- (G);
      \draw [det,->] (G) -- (I) node[midway,below,sloped] {$\mathrm{ext}$};
      \draw [det,->] (G) -- (P) node[midway,below,sloped] {$\mathrm{prune}$};
      \draw [det,->] (P) -- (V) node[midway,below,sloped] {$\mathrm{obs}$};
      \draw [det,->] (P) -- (O);
      \draw [det,->] (V) -- (W);
      \draw [stoch,->] (H) -- (G);
    \end{tikzpicture}
  }
  \caption{
    \label{fig:constellation}
    Relations among the various Markov processes discussed in the paper.
    Deterministic maps are indicated with solid arrows;
    random maps are shown as dashed arrows.
    All the maps shown commute.
    $\X_t$ is the \emph{population process}, a model of the dynamics of some system, which we take as a starting point.
    $\Hist_t$ is the \emph{history process}, which records the full history of $\X_t$.
    $\Inv_t$ is the \emph{inventory process}: at each time $t$, $\Inv_t$ is an inventory of all extant individuals in the population, each of which has a globally unique name.
    $\Gen_t$ is the \emph{genealogy process}, which captures the precise genealogical relationships among all individuals in $\Inv_t$, as well as among any samples that have been taken from the population.
    $\Vis_t$ is the \emph{visible genealogy process}, which is $\Gen_t$ pruned so that only relationships among samples remain.
    Finally $\Wch_i$ is the \emph{embedded chain of the visible genealogy process}, which is $\Vis_{s_i}$, $s_i$ being the time of the $i$-th sample.
    All of these processes can be obtained via deterministic procedures applied to the \emph{master process} $\Master_t$, as described in the text.
  }
\end{figure}

@@

\begin{lemma}\label{lemma:ell}
  Define time intervals $\Delta_{i}^{j}$ and $\Delta_{i}^{j,j'}$ as follows:
  \begin{equation*}
    \Delta_{i}^{j}\coloneqq\begin{cases}
      \halfopen{a_i^j,s_i^j}, & \Tset^j(\Och_i)\setminus\Tset^j(\Och_{i-1})\neq\emptyset, \\
      \emptyset, &\textrm{otherwise},
    \end{cases}
    \quad\textrm{and}\quad
    \Delta_{i}^{j,j'}\coloneqq\begin{cases}
      \halfopen{z_i^{j,j'},s_i^j}, & \Zset^{j,j'}(\Och_i)\setminus\Zset^{j,j'}(\Och_{i-1})\neq\emptyset, \\
      \emptyset, &\textrm{otherwise},
    \end{cases}
  \end{equation*}
  where $s_i^j=\Tset^j(\Och_i)\setminus\Tset^j(\Och_{i-1})$, $a_i^j=\Masterset^j(\Och_i)\setminus\Masterset^j(\Och_{i-1})$, and $z_i^{j,j'}=\Zset^{j,j'}(\Och_i)\setminus\Zset^{j,j'}(\Och_{i-1})$ are the corresponding sample, attachment times, respectively, of the $i$-th sample lineage in Deme $j$, and the deme-transition times from Deme $j$ to $j'$, respectively.
  Then, if it is understood that $\ell^j(t,\Och_0)=0$ for all $j$,
  \begin{equation*}
    \ell^j(t,\Och_i)=\ell^j(t,\Och_{i-1})+\indicator{\Delta_{i}^{j}}(t)-\sum_{j'}\left(\indicator{\Delta_{i}^{j,j'}}(t)-\indicator{\Delta_{i}^{j',j}}(t)\right), \qquad \forall i>0.
  \end{equation*}
  Trivially, $\ell(t,\Och_i)=\sum_j\ell^j(t,\Och_i)$.
\end{lemma}

\section{Results}
\label{sec:results}
The derivation of the probability density of $\Wch_i$ given $\Hist_{S_i}$ can be divided into two steps:
\begin{align}
\begin{split}
  P_{\Wch_i|\Hist}&=\sum_{\Och_i}\,P_{\Wch_i|\Och_i}P_{\Och_i|\Hist} \\
                  &=\sum_{\Och_i}\,P_{\Wch_i|\Och_i}\left(P_{\Och_1|\Hist}\prod_{q=2}^{i}P_{\Och_q|\Och_{q-1}}\right) 
\end{split}
\end{align}
% Trivially, $P_{\Och_1|\Hist}=1$ and $P_{\Wch_1|\Och_1}=1$ such that $P_{\Wch_1|\Hist}=1$.

@@

We define $\phi_{qk}^{j}\coloneqq\sum_{\omega\in\Omega_q}\dfrac{\indicator{\Inv_{t_{k-1}}^j(\omega)}(N_k(\omega))}{I^j(x_k-u_k)}$.

@@

All cases can be categorized into four groups based on the occurrence of population event within time $\halfopen{a_q,s_q}$: no population event or simultaneous events, a birth from Deme $j$ to $j'$, a sample of Deme $j'$, and a migration from Deme $j$ to $j'$ where $j\neq j'$. Within each group, we investigate $\phi_{qk}^{j''}$ by considering whether the corresponding population event was presented in the pruned genealogy and involving the $q$-th lineage sample at time $t_k$, given that the $q$-th lineage sample was in Deme $j''$ at time $t_k$. For $q>1$, we then have:

% 1) No. of total events in the population
% 2) No. of conditioned event in the population
% 3) possibilities that presented in the genealogy
% 4) probability of the exact sample involved

\begin{enumerate}[(a)]
  \item There was no population event within $\halfopen{a_q,s_q}$ or more than one population events occurring that involved the $q$-th sample:
    \begin{enumerate}
      \item if $t_k\notin\halfopen{a_q,s_q}$, or if $u_k\notin\Bset\cup\Sset\cup\Mset$, $\phi_{qk}^{j''}=1$;
      \item if $t_k\in\Cset(\Och_{q-1})\cup\Qset(\Och_{q-1})$, $\phi_{qk}^{j''}=1$.
    \end{enumerate}

  % M
  \item There was a migration event, that is, $u_k\in\Mset^{j,j'}$ within $\halfopen{a_q,s_q}$, given that it was not presented in the pre-existing genealogy $\Och_{q-1}$, \ie $t_k\notin\Zset^{j,j'}(\Och_{q-1})$: 
    \begin{enumerate}
      \item if this migration event exactly involved the $q$-th sample and was presented as a deme-transition event in $\Och_q$, \ie $t_k\in\Zset^{j,j'}(\Och_q)$, then $\phi_{qk}^{j''}=\delta_{j',j''}/\left((I_{k-1}^{j}-\ell_{q,k-1}^{j})(I_{k}^{j'}-\ell_{qk}^{j'})\right)$;
      \item otherwise, $t_k\notin\Zset^{j,j'}(\Och_q)$, then $\phi_{qk}^{j''}=1-\delta_{j',j''}/\left((I_{k-1}^{j}-\ell_{q,k-1}^{j})(I_{k}^{j'}-\ell_{qk}^{j'})\right)$.
    \end{enumerate}

  % G
  \item There was a sample event, that is $u_k\in\Sset^{j'}$ within $\halfopen{a_q,s_q}$:
    \begin{enumerate}
      \item if the lineage sampled at this event served as a direct ancestor to the $q$-th sample lineage, that is, $t_k=a_q\in\Lset^{j'}(\Och_{q-1})$, this indicates that among $I_k^{j'}-\ell_{qk}^{j'}$ unsampled individuals in deme $j'$ at $t_k$, the $q$-th sample lineage is the exact direct descent, then $\phi_{qk}^{j''}=\delta_{j',j''}/\left(I_k^{j'}-\ell_{qk}^{j'}\right)$;
      \item otherwise, $t_k\in(a_q,s_q)\setminus\Qset(\Och_{q-1})$, and $\phi_{qk}^{j''}=1-\delta_{j',j''}/\left(I_k^{j'}-\ell_{qk}^{j'}\right)$.
    \end{enumerate}

  % B, j=j'
  \item There was a birth event within the same deme, that is $u_k\in\Bset^{j,j'}$ and $\delta_{j,j'}=1$, but it was not shown in the genealogy yet because only one of the pre-existing $\ell_{qk}^{j'}$ lineages involved, \ie $t_k\notin\Cset^{j,j'}(\Och_{q-1})$:
    \begin{enumerate}
      % there were in total \binom{I}{2} combinations presented as births, while \binom{l}{2} is excluded because the birth was not presented in genealogy.
      \item if this birth event was presented as a coalescent event in $\Och_q$, \ie $t_k\in\Cset^{j,j'}(\Och_q)$, with probability $\ell_{qk}^{j'}/\left(\binom{I_k^{j'}}{2}-\binom{\ell_{qk}^{j'}}{2}\right)$, then the probability that the $q$-th sample attached to \emph{a certain} sample lineage in the genealogy is $1/\ell_{qk}^{j'}$, implying that the $q$-th sample lineage was exactly the other individual involved in the birth event, and eventually, $\phi_{qk}^{j''}=\left(\delta_{j,j'}\delta_{j',j''}\right)/\left(\binom{I_k^{j'}}{2}-\binom{\ell_{qk}^{j'}}{2}\right)$;
      \item otherwise, $t_k\notin\halfopen{a_q,s_q}$,  $\phi_{qk}^{j''}=1-\left(\delta_{j,j'}\delta_{j',j''}\ell_{qk}^{j'}\right)/\left(\binom{I_k^{j'}}{2}-\binom{\ell_{qk}^{j'}}{2}\right)$.
    \end{enumerate}

  % B, j!=j'
  \item There was a birth event between two different demes, that is $u_k\in\Bset^{j,j'}$ and $\delta_{j,j'}=0$:
    \begin{enumerate}
      % There were  l'*(I-l) possibilities of transitions
      \item and given that it was presented as a deme-transition event in the pre-existing genealogy, that is, $t_k\in\Zset^{j,j'}(\Och_{q-1})$:
        \begin{enumerate}
          % the genealogy has the exact l' lineage as the child, and the sample is the exact 1 as the parent
          % 1/[l'*(I-l)] * l' * 1/1
          \item if the $q$-th sample was from deme $j$ and is exactly the parent in the birth, that is, $t_k\in\Cset^{j,j'}(\Och_q)$ and $\delta_{j,j''}=1$, then $\phi_{qk}^{j''}=\delta_{j,j''}/\left(I_k^{j}-\ell_{qk}^{j}\right)$;
          \item otherwise, $t_k\in\Zset^{j,j'}(\Och_q)$,and $\phi_{qk}^{j''}=1-\delta_{j,j''}/\left(I_k^{j}-\ell_{qk}^{j}\right)$;
        \end{enumerate}
      % there were in total I*I' in the birth, while l*l' is excluded since it was not presented as coalescent, and (I-l)*l' are excluded which as presented as transition. == I I' - l l' - (I-l) l' = I (I'-l')
      \item or given that it was not presented in $\Och_{q-1}$, \ie $t_k\notin\Zset^{j,j'}(\Och_{q-1})\cup\Cset^{j,j'}(\Och_{q-1})$:
        \begin{enumerate}
          % the genealogy has l lineages as parent, and we choose one of them 1/l: 1/[I (I'-l')] * l * 1/l
          \item if the $q$-th sample was from deme $j'$ and is coalesced at $t_k$, \ie $t_k\in\Cset^{j,j'}(\Och_q)$ and $\delta_{j',j''}$, then $\phi_{qk}^{j''}=\delta_{j',j''}/\left(I_k^{j}(I_{k}^{j'}-\ell_{qk}^{j'})\right)$;
          % 
          \item if the $q$-th sample was from deme $j'$ and is presented as a deme-transition event in $\Och_q$, \ie $t_k\in\Zset^{j,j'}(\Och_q)$, then $\phi_{qk}^{j''}=\delta_{j',j''}/\left(I_k^{j}(I_{k}^{j'}-\ell_{qk}^{j'})\right)$;

          \item otherwise, $t_k\notin\Cset^{j,j'}(\Och_q)\cup\Zset^{j,j'}(\Och_q)$, and $\phi_{qk}^{j''}=1-\delta_{j',j''}/\left(I_{k}^{j'}-\ell_{qk}^{j'}\right)$;
        \end{enumerate}
    \end{enumerate}
\end{enumerate}

@@

    % deme-transition time can evolve to coalescent time, when the new sample is the parent of a "cross-deme" birth
    1, &\text{if}\ t_k\in\Cset(\Och_{q-1})\cup\Qset(\Och_{q-1}),\\[12pt]
    \frac{\delta_{j',j''}}{(I_{k-1}^j-\ell_{q,k-1}^j)(I_k^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\in\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Mset^{j,j'},\\[12pt]
    1-\frac{\delta_{j',j''}}{(I_{k-1}^j-\ell_{q,k-1}^{j})(I_k^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\notin\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Mset^{j,j'},\\[12pt]

    % sample
    \frac{\delta_{j',j''}}{I_k^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k=a_q\in\Lset^{j'}(\Och_{q-1})\ \text{and}\ u_k\in\Sset^{j'},\\[12pt]
    1-\frac{\delta_{j',j''}}{I_k^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k\in(a_q,s_q)\setminus\Qset(\Och_{q-1})\ \text{and}\ u_k\in\Sset^{j'},\\[12pt]

    % B, j=j'
    \frac{\delta_{j,j'}\delta_{j',j''}}{\binom{I_k^{j'}}{2}-\binom{\ell_{qk}^{j'}}{2}}, &\text{if}\ t_k=a_q\in\Cset^{j,j'}(\Och_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j=j',\\[12pt]
    1-\frac{\delta_{j,j'}\delta_{j',j''}\ell_{qk}^{j'}}{\binom{I_k^{j'}}{2}-\binom{\ell_{qk}^{j'}}{2}}, &\text{if}\ t_k\in(a_q,s_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j=j',\\[12pt]

    % B, j!=j'
    % in Z
    \frac{\delta_{j,j''}}{I_k^{j}-\ell_{qk}^{j}}, &\text{if}\ t_k\in\Zset^{j,j'}(\Och_{q-1}),\ t_k=a_q\in\Cset^{j,j'}(\Och_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]
    1-\frac{\delta_{j,j''}}{I_k^{j}-\ell_{qk}^{j}}, &\text{if}\ t_k\in\Zset^{j,j'}(\Och_{q-1}),\ t_k\in\Zset^{j,j'}(\Och_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]
    % not in Z
    \frac{\delta_{j',j''}}{I_k^j(I_{k}^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\notin\Zset^{j,j'}(\Och_{q-1}),\ t_k\in\Zset^{j,j'}(\Och_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]
    \frac{\delta_{j',j''}}{I_k^j(I_{k}^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\notin\Zset^{j,j'}(\Och_{q-1}),\ t_k=a_q\in\Cset^{j,j'}(\Och_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]
    1-\frac{\delta_{j',j''}}{I_{k}^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k\notin\Zset^{j,j'}(\Och_{q-1}),\ t_k\in(a_q,s_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]

    0, &\text{otherwise}.
  \end{cases}
\end{equation}

% To summarize, we have
% \begin{equation}\label{eq:qcases}
%   \phi_{qk}^{j''}=
%   \begin{cases}
%     1, &\text{if}\ t_k\notin\halfopen{a_q,s_q}\ \text{or}\ u_k\notin\Bset\cup\Sset\cup\Mset,\\
%     
%     % deme-transition time can evolve to coalescent time, when the new sample is the parent of a "cross-deme" birth
%     1, &\text{if}\ t_k\in\Cset(\Och_{q-1})\cup\Qset(\Och_{q-1}),\\[12pt]
%     \frac{\delta_{j',j''}}{(I_{k-1}^j-\ell_{q,k-1}^j)(I_k^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\in\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Mset^{j,j'},\\[12pt]
%     1-\frac{\delta_{j',j''}}{(I_{k-1}^j-\ell_{q,k-1}^{j})(I_k^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\notin\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Mset^{j,j'},\\[12pt]
%     
%     % sample
%     \frac{\delta_{j',j''}}{I_k^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k=a_q\in\Lset^{j'}(\Och_{q-1}),\\[12pt]
%     1-\frac{\delta_{j',j''}}{I_k^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k\in(a_q,s_q)\setminus\Qset(\Och_{q-1})\ \text{and}\ u_k\in\Sset^{j'},\\[12pt]
%     \frac{\delta_{j',j''}}{(I_k^j-\ell_{qk}^j)(I_k^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\in\Zset^{j,j'}(\Och_q)\setminus\Zset^{j,j'}(\Och_{q-1}),\ \text{and}\ u_k\in\Bset^{j,j'},\\[12pt]
%     
%     % B2C: either j''=j (in Z_{q-1} or j''=j' (not in Z_{q-1})
%     % the numerator = 1 when j'!=j and =2 when j'=j.
%     \frac{\delta_{j,j''}+\delta_{j',j''}}{I_k^{j}(I_k^{j'}-\delta_{j,j'})-\ell_{qk}^j(\ell_{qk}^{j'}-\delta_{j,j'})}, &\text{if}\ t_k=a_q\in\Cset^{j,j'}(\Och_q),\\[12pt]
%     1-\frac{(\delta_{j,j''}+\delta_{j',j''})\ \ell_{qk}^{j}\ \delta_{j',j''}}{I_k^{j}(I_k^{j'}-\delta_{j,j'})-\ell_{qk}^j(\ell_{qk}^{j'}-\delta_{j,j'})}-\frac{(1-\delta_{j,j'})\delta_{j',j''}}{I_k^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\notin\Zset^{j,j'}(\Och_{q-1})\cup\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Bset^{j,j'},\\[12pt]
%     1-\frac{\delta_{j,j''}}{I_k^{j}I_k^{j'}-\ell_{qk}^j\ell_{qk}^{j'}}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\in\Zset^{j,j'}(\Och_{q-1}),\ \text{and}\ u_k\in\Bset^{j,j'},\\[12pt]
%     0, &\text{otherwise}.
%   \end{cases}
% \end{equation}

@@

\begin{thm}\label{thm:prunehthm1}
  With the definitions as above, $P_{\Och_i|\Hist}(o|h) = \prod_{q=1}^{i}\prod_{k=1}^{K}\left(\sum_{j''=1}^{J}\phi_{qk}^{j''}\right)=\prod_{q=1}^{i}\prod_{k=1}^{K}\phi_{qk}^{j''_q}$, where $j''_q$ is the deme of the $q$-th sample lineage.
\end{thm}
We now exchange the order of multiplication over $q$ and $k$ then we can derive our \Cref{thm:prunethm2}.

\begin{thm}\label{thm:prunethm2}
  \begin{equation*}
    P_{\Prune_t|\Hist_t}(\Prune_t|h)=.
  \end{equation*}
\end{thm}
\begin{proof}
  Let $\Prune_t=\Och_i$ for some $i$.

  Last but not least, suppose $u_k\in\Bset^{j,j'}$.

\end{proof}
